\thetitle

### 2021-03-02 writing/mutate-benchmarks-figure.tex
editiroal voice thru 6

\begin{figure}
\begin{tabular}{p{1.75cm} | p{6cm} }
  {\bf  name} & {\bf description (author)}  \\

\hline
  \texttt{acquire} & Object-oriented board game implementation (M. Felleisen)  \\%[1em]

\hline
  \texttt{gregor} & Utilities for calendar dates (J. Zeppieri) \\%[1em]

\hline
  \texttt{kcfa} & Functional implementation of 2CFA for the lambda calculus (M. Might) \\%[1em]

\hline
  \texttt{quadT} & Converter from S-expression source code to PDF format (M. Butterick)\\%[1em]

\hline
  \texttt{quadU} & Converter from S-expression source code to PDF format  (B. Greenman) \\%[1em]

\hline
  \texttt{snake} & Functional implementation of the  Snake video game (D. Van Horn) \\%[1em]

\hline
  \texttt{suffixtree} & Algorithm for common longest subsequences between strings. (D. Yoo) \\%[1em]

\hline
  \texttt{synth} & Converter of notes and drum beats to WAV (V. St-Amour \& N. Toronto) \\%[1em]

\hline
  \texttt{take5} & Mixin-based card game simulator (M.Felleisen)  \\%[1em]

\hline
  \texttt{tetris} & Functional implementation of Tetris (D. Van Horn) \\%[1em]

\end{tabular}
  \caption{Benchmarks summary.}
  \label{table:benchmark-descriptions}
\end{figure}

>>>>

\begin{figure*}
\begin{tabular}{p{1.4cm} | p{6.9cm} | p{2.0cm} | r | r}
             {\bf  name} & {\bf description} & {\bf author} & {\bf loc} & {\bf mod.} \\ \hline
%% ---------------------------------------------------------------------------------------------------------------

\texttt{acquire} & object-oriented board game implementation      & M. Felleisen & 1941 & 9 \\ \hline

\texttt{gregor}  & utilities for calendar dates                   & J. Zeppieri  & 2336 & 13\\ \hline

\texttt{kcfa}    & functional implementation of 2CFA for LC       & M. Might & 328 & 7\\ \hline

\texttt{quadT}   & converter from S-expression source code to PDF & M. Butterick & 7396 & 14\\ \hline

\texttt{quadU}   & converter from S-expression source code to PDF & B. Greenman  & 7282 & 14 \\ \hline

\texttt{snake}   & functional implementation of the Snake game    & D. Van Horn & 182 & 8 \\ \hline

\texttt{synth}  & converter of notes and drum beats to WAV & V. St-Amour & 871 & 10 \\ \hline

\texttt{take5} & Mixin-based card game simulator & M.Felleisen & 465 & 8\\ \hline

\texttt{tetris} & Functional implementation of Tetris & D. Van Horn &   280 & 9 \\ \hline

\texttt{suffix\-tree} & algorithm for common longest subsequences between strings. & D. Yoo & 1500 & 6 \\

\end{tabular}

\caption{Benchmarks summary.} \label{table:benchmark-descriptions}

\end{figure*}

### 2021-03-02 writing/mutate-benchmarks.tex
editiroal voice thru 6

Since this is the first empirical attempt to evaluate blame in gradual
typing, there is no standard benchmark suite that we can use off-the-self.
In addition, there is no exsitng benchmark suite of gradually typed
programs with type-level mistakes that we could
repurpose.\footnote{Recently ~\citet{cc-oopsla-20} created a
Reticulated Python program collection to evaluate their technique for
fixing mistakes in type annotations; unfortunately their collection does
not come with type-level mistakes in the code proper.} In fact, the bar
for testing the rational programmer is much higher than the existence of a
collection of programs with type-level errors. Testing the rational
programmer demands a currated collection of programs that take advantage
of the sophisticated features of gradually typed languages such as Typed
Racket.

>>>>

%% -----------------------------------------------------------------------------

Since this is the first empirical attempt to evaluate blame in gradual typing,
there is no standard benchmark suite that we can use off-the-self.  In addition,
there is no exsitng benchmark suite of gradually typed programs with type-level
mistakes.\footnote{\citet{cc-oopsla-20} created a collection of Reticulated
Python programs to evaluate their technique of fixing mistakes in type
annotations. Their collection does not come with type-level mistakes in the code
itself.} In fact, the bar for testing the rational programmer is much higher
than the existence of a collection of programs with type-level errors. Testing
the rational programmer demands a currated collection of programs that take
advantage of the sophisticated features of a gradually typed language such as
Typed Racket.

### 2021-03-02 writing/mutate-interesting.tex
bring back interesting=plot
editiroal voice thru 6

Let us consider the three conditions in order:
\begin{enumerate}

@@

%\begin{figure*}
%  \centering
%  \includegraphics[scale=0.35]{./plots/mutant-breakdown}
%  \caption{Interesting debugging scenarios produced by our mutators. Counts are cut off at 50.}
%  \label{fig:mutant-breakdown}
%\end{figure*}

>>>>

Here is the rationale for these three conditions:
\begin{enumerate}

@@

\begin{figure*} \footnotesize
 \centering
 \includegraphics[scale=0.33]{./plots/mutant-breakdown}
  Counts are cut off at 50.

\caption{The mutators produce Interesting debugging scenarios across all benchmarks.} 
 \label{fig:mutant-breakdown}
\end{figure*}

The definition of interesting debugging scenarios creates a powerful filter. All
together, the listed mutators produce 16,800 mutants with at least one
interesting debugging scenario across all benchmarks; see
figure~\ref{fig:mutant-breakdown} for an overview. Broken down by benchmark, the
mutators produce at least 40 interesting scenarios for every benchmark, and
these scenarios originate from at least four different mutators per benchmark.
Thus, the mutators result in a sizable and diverse population of scenarios for
every benchmark.  Furthermore, every mutator contributes scenarios to at least
one benchmark.  Some mutators apply only to a few benchmarks, because they
target rather specific features; for instance, the class-focused mutators are
mainly effective in a program that makes extensive use of object-oriented
features.

### 2021-03-02 writing/mutate-mutators.tex
editiroal voice thru 6

Figure~\ref{table:mutation-ops} describes 16 mutators that satisfy this
requirement.  For each mutator, the figure provides a short description and an
example. As the explanation indicates, some specialize or
generalize~\citet{lksfd-popl-2020}'s mutators, which in turn are borrowed from
the vast literature on mutation testing~\cite{jia2011analysis}.\footnote{Many
of~\citet{lksfd-popl-2020} mutators do not satisfy the constraints. For
example, the relational mutator does not lead to type-level errors that Typed
Racket's type system can detect, so it has been eliminated. Similarly, the
original arithmetic mutator does not reliably lead to type errors, so it has
been adapted to only make operator swaps that can affect the type of the
result. Changing a `/` to `*` does not affect the type of an expression, but
changing `*` to `/` may, because division by zero is detectable by Typed
Racket's type system.}  Only two are directly inherited; many mutators are
brand new. For the latter the authors relied on their decades-long experience
of making type-level mistakes in Typed Racket, some of which take a non-trivial
effort to debug.

>>>>

Figure~\ref{table:mutation-ops} describes 16 mutators that satisfy these
constraints. As the explanation indicates, some specialize or
generalize~\citet{lksfd-popl-2020}'s mutators, which in turn are borrowed from
the vast literature on mutation testing~\cite{jia2011analysis}.\footnote{Many
of~\citet{lksfd-popl-2020} mutators do not satisfy the constraints and had to
be eliminated. For example, the relational mutator does not lead to type-level
errors. Similarly, the arithmetic mutator does not reliably lead to type
errors; it had to be adapted to make only operator swaps that can affect the
type of the result, e.g., changing `*` to `/`.}  Only two are directly
inherited; many mutators are brand new. For the latter the authors relied on
their decades-long experience of making type-level mistakes in Typed Racket,
some of which take a non-trivial effort to debug.
